---
title: "team.Rmd"
author: "Elaine OB. and Andy M."
date: "4/19/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(maps)
library(plotly)
library(gridExtra)
```

Part(1)

1. Are there some days of the week where more accidents happen than on others (use variable DAY_WEEK)?  

##### Friday, Saturday, and Sunday are the most popular days in which accidents occur. 

2. What time of the day do accidents happen (use variable HOUR)?

##### The time of day in which accidents happen is between 5pm and 12am


3. What is the number of accidents with at least one drunk driver (use variable DRUNK_DR)?


##### 

```{r}

accident <-  read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)

person <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)

```


```{r}
names(accident)
names(person)

```


```{r}
#summary(Accident$DAY_WEEK)
#summary(Accident$HOUR)
#Changing labels was not thought up by our team.  We stumbled across this idea @ http://www.sthda.com/english/wiki/ggplot2-axis-ticks-a-guide-to-customize-tick-marks-and-labels

Best_Days_To_Get_Into_Accident <- accident %>% ggplot(aes(x = DAY_WEEK))+
  geom_histogram(bins = 7, binwidth = .5)+ scale_x_continuous(breaks = seq(1, 7), labels = c("Sun", "Mon", "Tues", "Wed", "Thurs", "Fri", "Sat"))+
  ggtitle("Popular Days Accidents Occur")+ xlab("Days of Week")+ ylab("Number of Accidents")

Best_Times_To_Get_Into_Accident <- accident %>% filter(HOUR <= 24) %>% ggplot(aes(x = HOUR))+
  geom_histogram(bins = 24, binwidth = .5)+ scale_x_continuous(breaks = seq(0, 23), labels = c("12am", "1am", "2am", "3am", "4am", "5am", "6am", "7am", "8am", "9am", "10am", "11am", "12pm", "1pm", "2pm", "3pm", "4pm", "5pm", "6pm", "7pm", "8pm", "9pm", "10pm", "11pm"))+ theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))+
  ggtitle("Popular Hours Accidents Occur throught a Day")+ xlab("Hour of Day")+ ylab("Number of Accidents")

grid.arrange(Best_Days_To_Get_Into_Accident, Best_Times_To_Get_Into_Accident, nrow = 2)

```


```{r}

accident %>% select(DRUNK_DR) %>% filter(DRUNK_DR >= 1) %>% summarize(n())

```
Part(2)

1. Load the person table. Identify drivers (PER_TYP == 1, see fars manual) and subset on them.

2. Join accident and driver table (work out which variable(s) to use)

3. Tally the number of accidents by day of the week (DAY_WEEK), hour of the day (HOUR) and gender (SEX). Visualize the results!


```{r}

driver_1 <- person %>% subset(PER_TYP == 1)

Joined_Data_ACCIDENTdrivers <- full_join(driver_1, accident)

Tally_NumberOfAccidents <- Joined_Data_ACCIDENTdrivers %>% group_by(DAY_WEEK) %>% summarize(number = n())
Tally_NumberOfAccidents

Tally_HourOfDay <- Joined_Data_ACCIDENTdrivers %>% group_by(HOUR) %>% filter(HOUR <= 24) %>% summarize(number = n())
Tally_HourOfDay


#SEX is an interesting variable. Male(1) and Female(2). Looking at the FARS Analytical Report we see there are drivers with unidentified SEX: Not Reported(8) or Unknown(9)
Tally_Sex_MF <- Joined_Data_ACCIDENTdrivers %>% group_by(SEX) %>% filter(SEX == 1 | SEX == 2) %>% summarize(number = n())
Tally_Sex_MF

Tally_Sex_NotReported <- Joined_Data_ACCIDENTdrivers %>% group_by(SEX) %>% filter(SEX == 8) %>% summarize(number = n())
Tally_Sex_NotReported

Tally_Sex_Unkown <- Joined_Data_ACCIDENTdrivers %>% group_by(SEX) %>% filter(SEX == 9) %>% summarize(number = n())
Tally_Sex_Unkown

```


```{r}

Tally_NumberOfAccidents %>% ggplot(aes(x = DAY_WEEK, y = number ))+ geom_bar(stat = "identity") + geom_text(aes(label = c("Sun", "Mon", "Tues", "Wed", "Thurs", "Fri", "Sat"), vjust = -.2))+ theme(axis.text.x = element_blank(), axis.text.y = element_text(face = "bold", color = "black"))+
  ggtitle("Tally - number of accidents by day ")+ xlab("Days of Week")+ ylab("Number of Accidents")

```


```{r}

Tally_HourOfDay %>% ggplot(aes(x = HOUR, y = number ))+ geom_bar(stat = "identity")+
  geom_text(aes(label = c("12am", "1am", "2am", "3am", "4am", "5am", "6am", "7am", "8am", "9am", "10am", "11am", "12pm", "1pm", "2pm", "3pm", "4pm", "5pm", "6pm", "7pm", "8pm", "9pm", "10pm", "11pm"), angle = 90), hjust = 2 )+
   ggtitle("Tally - number of accidents by hour in a Day ")+ xlab("Hour of a Day")+ ylab("Number of Accidents")

```


```{r}
Tally_Sex_MF %>% ggplot(aes(x = SEX, y = number))+ geom_bar(stat = "identity")+
  geom_text(aes(label = c("Male", "Female"), vjust = 3.7), size = 6)+
  ggtitle("Tally - Male and Female ")+ xlab("Sex Identified")+ ylab("Number of Accidents")
  

```
Part(3)

1. Plot a choropleth map of accidents on a county level

2. Read the description for the STATE and COUNTY columns in the FARS manual.

3. The state & county codes are Geographic Locator Codes (GLCs) from the General Services Administration’s (GSA) publication

4. Use readxl::read_xlsx to read in the GLCs

5. Visualize the number of accidents in each county using a choropleth map. To do this, you need to join the accident data, GLCs, and appropriate map data.

6. Can you find seasonal trends in accident occurrences in different states? Is summer or winter more dangerous? Why?



```{r}

FRPP_GLC <- readxl::read_xlsx("FRPP GLC United States.xlsx")

colnames(FRPP_GLC) <- FRPP_GLC[1,]
FRPP_GLC <- FRPP_GLC[-1,]

county <- map_data('county')

FRPP_GLC_rename <- FRPP_GLC %>% mutate(STATE = as.integer(`State Code`), COUNTY = as.integer(`County Code`))

Map_Join <- full_join(accident, FRPP_GLC_rename) %>% select("STATE", "COUNTY", "County Name", "State Name", "ST_CASE") %>% mutate(region = tolower(`State Name`), subregion = tolower(`County Name`)) %>% select("STATE", "COUNTY", "region", "subregion") %>% group_by(region, subregion) %>% summarize(number = n())   #Code Sitation   https://stackoverflow.com/questions/45357045/convert-a-column-in-r-data-frame-to-lower-case

Map_Join_df <- full_join(county, Map_Join, by = c("region", "subregion"))

Map_Join_df %>% ggplot(aes(x=long, y=lat))+
  geom_polygon(aes(group= group))+ 
  geom_point(aes(x=LONGITUD, y=LATITUDE), accident %>% 
               filter(LONGITUD < 0 & LATITUDE > -130),size=0.02, color='white')


Map_Join_df %>% ggplot(aes(x = long, y = lat, fill = log(number)))+ geom_polygon(aes(group = group))+ ggtitle("Visualization of accidents by County")+ xlab("Longitude")+ ylab("Latitude") 

```